---
title: "R Notebook"
output: html_notebook
---

```{r}
library(ggplot2)
library(magrittr)
library(tidyverse)
library(marima)
library(stats)
```

# Read data

```{r}
data <- read.csv(file = 'DataAssignment3.csv')
head(data)
```



## Plot data

```{r}
(plot1 <- ggplot(data = data)+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_rect(data=data,aes(xmin = Date, xmax = Date+1,
                 ymin = -Inf, ymax = Inf ,fill=Government),alpha=0.55)+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_line(aes(x=Date,y=NumberWords)))

```


## Transform data

Use log and second order differencing on NumberWords to get transformed column NumberWordsTrans

```{r}
diff_order <- 2

data1 <- data %>%
  mutate(NumberWordsTrans = data %>% .$NumberWords %>% log() %>%  diff(differences = diff_order) %>% c(rep(NA, diff_order), .)) %>% 
  mutate(Conservative = ifelse(Government == "Conservative", 1,0)) %>% 
  mutate(Independent = ifelse(Government == "Independent", 1,0)) %>%
  mutate("Social Democrats" = ifelse(Government == "Social Democrats", 1,0)) %>%
  mutate("Social Liberals" = ifelse(Government == "Social Liberals", 1,0)) %>%
  mutate("Venstre" = ifelse(Government == "Venstre", 1,0))
data1


```

## Plot transformed data

```{r}
(plot <- ggplot(data = data1)+
  geom_point(aes(x=Date,y=NumberWordsTrans))+
  geom_rect(data=data,aes(xmin = Date, xmax = Date+1,
                 ymin = -Inf, ymax = Inf,fill=Government),alpha=0.55)+
  geom_point(aes(x=Date,y=NumberWordsTrans))+
  geom_line(aes(x=Date,y=NumberWordsTrans)))
```


# MARIMA

Fit MARIMA model and forecast

## Own differencing

### Fit model

```{r}
ar = c(1)
ma = c(1)
data2 <- data1[, c(1:2, 4:9)] # manually remove government column

Model1 <- define.model(kvar = ncol(data2), rem.var = c(1,2), ma=ma, ar = ar, reg.var = c(4:8))
Marima1 <- marima(t(data2), ar.pattern=Model1$ar.pattern, ma.pattern=Model1$ma.pattern, Plot="log.det", penalty=0)

# marima_model stats
print("MA params:")
short.form(Marima1$ma.estimates, leading = F)
print("MA p-values")
Marima1$ma.pvalues
print("AR params:")
short.form(Marima1$ar.estimates, leading = F)
print("AR p-values")
Marima1$ar.pvalues


```


### Forecasting

#### Add forecasting to data

```{r}
data3 <- data2 %>% 
  mutate("forecast" = arma.forecast(data2, nstart=92, nstep=10, marima=Marima1, check = F)$forecasts[3,] ) %>% 
  mutate("forecast_inv_trans" = 
           forecast[3:102] %>% # first two are used for diffinv conditions
           diffinv(differences = 2, xi = log(data2$NumberWords[1:2])) %>% 
           exp()) %>% 
  mutate(pred_var = c(rep(0, 12), forecasts$pred.var[8,8,])) %>% 
  mutate("forecast_inv_trans_lower" = 
           (forecast - sqrt(pred_var)*1.96)[3:102] %>% 
           diffinv(differences = 2, xi = log(data2$NumberWords[1:2])) %>% 
           exp()) %>% 
  mutate("forecast_inv_trans_upper" = 
           (forecast + sqrt(pred_var)*1.96)[3:102] %>% 
           diffinv(differences = 2, xi = log(data2$NumberWords[1:2])) %>% 
           exp()) %>% 
  mutate("residuals" = NumberWords-forecast_inv_trans) 
  # mutate("restore_org" = 
  #         NumberWordsTrans[3:102] %>% 
  #          diffinv(differences = 2, xi = log(data2$NumberWords[1:2])) %>% 
  #          exp())
  # select(NumberWords, restore_org, NumberWordsTrans)

data3

```


#### Plot forecast

```{r}
(plot <- ggplot(data = data3)+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_rect(data=data,aes(xmin = Date, xmax = Date+1,
                 ymin = -Inf, ymax = Inf,fill=Government),alpha=0.55)+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_line(aes(x=Date,y=NumberWords)) +
  geom_point(aes(x = Date, y = forecast_inv_trans), shape = 2)
  # geom_point(aes(x = Date, y = forecast_inv_trans_lower), shape = 2) + # lower prediction interval
  # geom_point(aes(x = Date, y = forecast_inv_trans_upper), shape = 2) # upper prediction interval
  )
```


## Differencing with define.dif


```{r}
data4 <- data2[-3]
data4["LogNumberWords"] <- log(data4$NumberWords)
differences <- matrix(c(8,1,8,1), nrow = 2)
dy <- define.dif(t(data4), differences)

ar <- c(1)
ma = c(1)
Model2 <- define.model(kvar = ncol(data4), rem.var = c(1,2), ma=ma, ar = ar, reg.var = c(3:7))
Marima2 <- marima(t(data2), ar.pattern=Model2$ar.pattern, ma.pattern=Model2$ma.pattern, Plot="log.det", penalty=.5)

# marima_model
print("MA params:")
short.form(Marima2$ma.estimates, leading = F)
print("MA p-values")
Marima2$ma.pvalues
print("AR params:")
short.form(Marima2$ar.estimates, leading = F)
print("AR p-values")
Marima2$ar.pvalues

```



### Forecasting

#### Add forecast to data


```{r}
data5 <- data4 %>% 
  mutate(forecast = arma.forecast(data2, marima = Marima2, nstart = 12, nstep = 90, dif.poly = dy$dif.poly)$forecasts[8,]) %>% 
  mutate(forecast_invtrans = forecast %>% exp())

data5
```

plot forecast

```{r}
(plot1 + 
   geom_point(data = data5, aes(x = Date, y = forecast_invtrans), shape = 2)
   )
```

# With functions


```{r}
data <- read.csv(file = 'DataAssignment3.csv')
```

```{r echo=TRUE}
trans <- function(data, diff_order = 2, trns = 1){  # data is a column
  if (trns == 1){
    data %>% log() %>%  diff(differences = diff_order) %>% c(rep(NA, diff_order), .)
  } else if (trns == 2){
    data %>%  diff(differences = diff_order) %>% log() %>% c(rep(NA, diff_order), .)
  }
  
  
}

inv_trans <- function(data, org_data, diff_order = 2, trns = 1){ # data is a column
  if (trns == 1){
    data %>% 
      .[(1+diff_order):length(data)] %>% 
      diffinv(differences = diff_order, xi = log(org_data[1:diff_order])) %>% 
      exp()
      
  } else if (trns == 2){
    data %>% 
      .[(1+diff_order):length(data)] %>% 
      exp() %>% 
      diffinv(differences = diff_order, xi = org_data[1:diff_order])
  } 
  
}

transform_data <- function(data, trns = 1, diff_order = 2){
  data %>%
  mutate(NumberWordsTrans = data %>% .$NumberWords %>% trans(trns = trns, diff_order = diff_order)) %>% 
  mutate(Conservative = ifelse(Government == "Conservative", 1,0)) %>% 
  mutate(Independent = ifelse(Government == "Independent", 1,0)) %>%
  mutate("Social Democrats" = ifelse(Government == "Social Democrats", 1,0)) %>%
  mutate("Social Liberals" = ifelse(Government == "Social Liberals", 1,0)) %>%
  mutate("Venstre" = ifelse(Government == "Venstre", 1,0)) %>% 
  select(-Government) #%>% 
  # mutate("org" = inv_trans(.$NumberWordsTrans, .$NumberWords))
}

add_forecast <- function(data, n_start = 102, n_step = 0, ar = c(1), ma = c(1), trns = 1, diff_order = 2, future_gov = "Social Democrats"){
  Model <- define.model(kvar = ncol(data), rem.var = c(1,2), ma=ma, ar = ar, reg.var = c(4:8))
  Marima <- marima(t(data), means = 1, ar.pattern=Model$ar.pattern, ma.pattern=Model$ma.pattern, Plot="log.det", penalty=0)
  
  ("MA params:")
  (short.form(Marima$ma.estimates, leading = F))
  ("MA p-values")
  (Marima$ma.pvalues)
  ("AR params:")
  (short.form(Marima$ar.estimates, leading = F))
  ("AR p-values")
  (Marima$ar.pvalues)
  
  
  return(Marima)
  
  
  n_future <- n_step + n_start - dim(data)[1]
  if (n_future > 0){
    data <- data %>% #.[dim(.)[1], ] %>% 
      bind_rows(
        data.frame(
          "Date" = c(2021:(2020+n_future)),
          "NumberWords" = NaN,
          "NumberWordsTrans" = NaN,
          "Conservative" = ifelse(future_gov == "Conservative", 1,0),
          "Independent" = ifelse(future_gov == "Independent", 1,0),
          "Social Democrats" = ifelse(future_gov == "Social Democrats", 1,0),
          'Social Liberals' = ifelse(future_gov == "Social Liberals", 1,0),
          "Venstre" = ifelse(future_gov == "Venstre", 1,0)
        ) %>% 
          rename("Social Democrats" = "Social.Democrats") %>% 
          rename('Social Liberals' = 'Social.Liberals')
      )
  }

  forecasts <- arma.forecast(data, nstart=n_start, nstep=n_step, marima=Marima, check = F)

  data %>%
    mutate("forecast" = forecasts$forecasts[3,] ) %>%
    mutate("forecast_inv_trans" = inv_trans(.$forecast, .$NumberWords)) %>%
    mutate(pred_var = c(rep(0, n_start), forecasts$pred.var[8,8,])) %>%
    mutate("forecast_inv_trans_lower" =
             (forecast - sqrt(pred_var)*1.96) %>%
             inv_trans(data$NumberWords, diff_order, trns)) %>%
    mutate("forecast_inv_trans_upper" =
             (forecast + sqrt(pred_var)*1.96) %>%
             inv_trans(data$NumberWords, diff_order, trns)) %>%
    mutate("residuals" = NumberWords-forecast_inv_trans)
}

# data %>% transform_data(trns = 1) %>% mutate(org = inv_trans(NumberWordsTrans, NumberWords, trns = 1))
data %>% transform_data() %>% add_forecast(n_start = 102, n_step = 10, ar = 1, ma = 1, diff_order = 1, trns = 1)
```


## Plots
```{r}
(plot <- data %>% transform_data() %>% add_forecast(n_start = 102, n_step = 10, ar = 1, ma = 1, diff_order = 1, trns = 1) %>%  ggplot()+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_rect(data=data,aes(xmin = Date, xmax = Date+1,
                 ymin = -Inf, ymax = Inf,fill=Government),alpha=0.55)+
  geom_point(aes(x=Date,y=NumberWords))+
  geom_line(aes(x=Date,y=NumberWords)) 
)

(plot +
  geom_line(aes(x = Date, y = forecast_inv_trans), color = "blue") +
  geom_line(aes(x = Date, y = forecast_inv_trans_lower), color = "red") + # lower prediction interval
  geom_line(aes(x = Date, y = forecast_inv_trans_upper), color = "red")  # upper prediction interval
)
# ggplot(data = data3) + 
#   geom_point(aes(x = Date, y = residuals))
```









```{r}

```



































